Programming Distributed Control Planes

### 2016-01-15 paper/background.tex
a pass

We focus on BGP as it is the most common and flexible way to implement distributed control planes today. Before delving into the details of our work, we provide here relevant background on BGP.

BGP is a path vector protocol that connects autonomous systems (ASes). An AS has one or more routers controlled by the same administrative entity. ASes exchange routing announcements with their neighbors. Each announcement has a destination IP prefix and some attributes (see below), and it indicates that the sending AS is willing to carry traffic to that prefix from the receiving AS. (Traffic flows in the opposite direction, from announcement receivers to senders.)

When an AS receives an announcement from a neighbor, it is processed by import filters that may drop the announcement or modify some attributes. If multiple announcements for the same prefix survive import filters, the router selects the best one based on local policy, expressed in terms of route attributes. This route is then used to send traffic to the destination. It is also advertised to the neighbors, after passing them through neighbor-specific export filters that may stop the announcement or modify some attributes.

A few route attributes are notable. The AS-path attribute accompanies all announcements and it has the ordered list of ASes that the announcement has traversed thus far. While AS-path attribute has a global meaning, some attributes are meaningful only within an AS or between neighboring ASes.  Two such attributes are communities and multi-exit discriminator (MED). Communities are opaque, string attributes that ASes use to color incoming routes on different criteria (e.g., ``entered on West Coast"). They can also be used to signal to neighbors about how the announcement should be handled (e.g., do not export it further). MEDs are used when an AS has more than one link to a neighboring AS. They are numeric values that signal to the receiving AS which link the sender prefers to receive traffic on.

The route selection process assigns a {\em local preference} to each route that survives the import filters. Routes with higher local preference are preferred. For routes with equal preference, other factors such as AS path length, MEDs, internal routing cost, are considered in order. Because they are considered first in the selection process, they This preference can be based on any route attribute. It can also be based on the sending neighbor, which is a common practice.

 , or other existing attributes.  Local preference is an internal attribute that is not carried beteeen the first criterion in selecting the best route and higher numbers are assigned to preferred routes. BGP conside

aggregation

>>>>

Our work focuses on BGP as it is the most common and flexible way to implement distributed control planes today. In this section, we provide relevant background on BGP.

BGP is a path-vector routing protocol that connects autonomous systems (ASes). An AS has one or more routers managed by the same administrative entity. ASes exchange routing announcements with their neighboring ASes. Each announcement has a destination IP prefix and some attributes (see below), and it indicates that the sending AS is willing to carry traffic to that prefix from the receiving AS. Traffic flows in the opposite direction, from announcement receivers to senders.

When a route announcement is received by an AS, it is processed by custom import filters that may drop the announcement or modify some attributes. If multiple announcements for the same prefix survive import filters, the router selects the best one based on local policy (see below). This route is then used to send traffic to the destination. It is also advertised to the neighbors, after processing through neighbor-specific export filters that may stop the announcement or modify some attributes.

All routing announcements are accompanied by an AS-path attribute that reflects the sequence of ASes that the announcement has traversed thus far. While the AS-path attribute has a global meaning, some attributes are meaningful only within an AS or between neighboring ASes.  One such attribute is a list of community strings. ASes use such strings to color routes on different criteria (e.g., ``entered on West Coast") and then use the color later in the routing process.  Communities are also used to signal to neighbors how they should handle an announcement (e.g., do not export it further). Another non-global attribute is the multi-exit discriminator (MED). It is used when an AS has more than one router-level link to a neighboring AS.  Its (numeric) values signal to the neighbor how this AS prefers to receive traffic among those links.

The route selection process assigns a {\em local preference} to each route that survives the import filters. Routes with higher local preference are preferred. Among routes with the same local preference, other factors such as AS path length, MEDs, internal routing cost, are considered in order. Because it is considered first during route selection, local preference is highly influential, and ASes may assign this preference based on any aspect of the route. A common practice is to assign it based on the commercial relationship with the neighbor. For instance, an AS may prefer in order customer ASes (which pay money), peer ASes (with free exchange of traffic), and provider ASes (which charge money for traffic).

The combination of arbitrary import and export filters and route selection policies at individual routers make BGP a highly flexible routing protocol. That flexibility, however, comes at the cost of it being difficult to configure correctly, as we explain next.

### 2016-01-15 paper/motivation.tex
a pass

We motivate our work by describing the current practice in network configuration and its shortcomings.
In traditional networks with distributed control planes, the operators’ goal is to generate the configuration of individual devices. Based on its configuration, the device exchanges information with other (e.g., links costs or best known paths) devices, filters and ranks incoming information, and decides how to forward packets to various destinations.

Device configuration languages are low-level and indirect. For instance, instead of allowing operators to express directly the paths they want through the network, they require operators to specify link costs that result in those paths; instead of allowing operators to express directly the types of traffic to not carry through the network, they require operators to select and program an appropriate filtering mechanism (e.g., route maps, null routing,  access control lists) and instantiate it on topologically appropriate devices; instead of allowing operators to directly specify that they prefer to send traffic via customer, peer, and provider networks, in that order, they require operators to program BGP local preferences at each router and ensure that the preferences are consistent across routers.

@@

To reduce such problems, some networks use a template-based approach. Configuration templates abstract certain constants into variables (e.g., instead of concrete interface IP address, they may contain a variable {\small \sf{\$$Rtr1\_Int1\_IP$)}} and may use a device vendor-neutral syntax. Operators manually generate the templates, and then use tools to translate them into device configuration, by replacing variables with appropriate constants using a database of network information.

@@

While templates avoids some pitfalls of the fully manual approach, they too are far from ideal. The fundamental issue is the semantic mismatch between desirable policies and the level of abstraction that templates offer. While desirable policies are network-wide (e.g., prefer customer networks, or link $A$ should serve as backup for link $B$), templates are device-level (e.g., BGP local preferences or link costs to use). Operators must still manually decompose network-wide policies into device-level policies that can produce the desired network-wide behavior.

This decomposition is not always straightforward and ensuring policy-compliance can be hard, especially in the face of failures. We illustrate this point using two examples, based on policies that we have seen in practice.

\subsection{OLD -- Overview}
\todo{this subsection can be deleted once we have captured everything in it}

Sources of bugs we fix. (It seems like the main advantage of our approach is not just fixing bugs though - it is easily describing high-level intention).
Perhaps the best way to do this is by going through a bunch of examples in section 2 and showing how you could easily introduce bugs:

\begin{itemize}
	\item Out of sync, or copy paste errors due to replicated configs (never an issue due to centralized control)
	\item Correct filtering to ensure no undesired traffic can flow through the network (e.g., best practices, like an AS should filter customers for their prefix, are implemented automatically )
	\item Failures can easily lead to unexpected behavior (e.g., a datacenter failure scenario w/instability.)
	\item Trying to do anything interesting, like get backups correct is difficult (e.g., setting up aggregation wrong)
	\item Related to the last point, things like aggregation can introduce black holes. We have the information needed to prevent this
\end{itemize}

Possible Examples:
\begin{itemize}
	\item Basic datacenter with spine preference
	\item Simple AS that prefers customers over peers over providers
	\item Combined internal backup routing with preference based entrance into the network using aggregation.
	\item Something like cold potato routing
\end{itemize}

>>>>

We motivate our work by describing the current practice in network configuration. In traditional networks with distributed control planes, the operators' goal is to generate the configuration of individual devices based on their desired policy. This configuration dictates the behavior of a device, how it exchanges routing information with neighbors and how it filters and ranks that information. The collective behavior of the devices should implement the desired policy.

Device configuration languages are low-level and indirect. For instance, instead of allowing operators to express directly the paths they want through the network, they require operators to specify metrics that result in those paths; instead of allowing operators to express directly the types of traffic to not carry through the network, they require operators to select and program an appropriate filtering mechanism (e.g., BGP import or export filters, null routing,  access control lists) and instantiate it on topologically appropriate devices; instead of allowing operators to directly specify that they prefer BGP neighbors in a certain order, they require operators to program local preferences and multi-exit discriminators (MED) at each router and ensure that the numbers are consistent across routers.

@@

To reduce such problems, some networks use a template-based approach. Configuration templates abstract certain constants into variables (e.g., instead of concrete community value, they may contain a variable {\small \sf{\$$BadNeighbor$)}} and may use a device vendor-neutral syntax. Operators manually generate the templates and use tools to translate them into device configuration, by replacing variables with appropriate constants using a database of network information.

@@

While templates avoids some pitfalls of the fully manual approach, they too are far from ideal. The fundamental issue is the semantic mismatch between desired policies and the level of abstraction of templates. While many policies are network-wide (e.g., prefer customer networks, or never announce a route to a certain prefix to external neighbors), templates are device-level. Operators must still manually decompose network-wide policies into device-level policies that can produce the desired network-wide behavior.
This decomposition is not always straightforward and ensuring policy-compliance can be hard, especially in the face of failures. We illustrate this point using two examples based on policies that we have seen in practice.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/example1}
\caption{Creating router-level policies is difficult.}
\label{fig:example1}
\end{figure}

Consider the backbone network in Figure~\ref{fig:example1}. It has three neighbors, a customer $Cust$ , a peer $Peer$, and a provider $Prov$. The policy of this network is shown on the right. Per commercial relationships, it prefers the neighbors in a certain order ($P1$) and does not want to act as a transit between Peer and Prov ($P2$). It prefers to exchange traffic with Cust over $R1$ rather than $R2$ because $R2$ is a more expensive link ($P3$). To guard against another AS "hijacking" prefixes owned by Cust, it only sends traffic to those prefixes if Cust is on the AS path ($P4$). Finally, to guard against Cust accidentally becoming a transit for Prov, it does not use Cust for traffic that will later traverse Prov ($P5$).

To correctly implement this network's policy, the operators must compute and assign local preferences such that preferences at Cust-facing interfaces $>$ Peer-facing interfaces $>$ Prov-facing interfaces. At the same time, the preference at $R2$'s Cust-facing interface should be lower than that at $R1$, but higher than that at Prov-facing interface. To fully realize $P2$, MEDs will have to be appropriately configured as well. To implement $P3$ and $P4$, the operators may assign communities that indicate where a certain routing announcement entered the network. Then, R4 must not announce to $Peer$ routes that have communities that correspond to the R2-prov link but must announce communities for the $R2$-Cust and $R1$-Cust links. Finally, to implement $P4$ and $P5$, the operators will have to compute and configure appropriate prefix- and AS-path-based filters at each router.

We can now see that devising correct configuration for real, larger networks will quickly become a nightmare. These networks will have many neighbors across multiple classes of commercial relationships, differing numbers of links per neighbor, along with several region-, neighbor- or prefix-based exceptions to the default policy. Templates only help to certain extent, e.g., by keeping preference and community values consistent across routers. The operators must still do much of the conceptually difficult work manually, in coming up with the templates themselves. \sysname enables a direct expression of the policies of such networks and compiles them automatically to individual router configurations.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/example2}
\caption{Policy-compliance under failures is difficult.}
\label{fig:example2}
\end{figure}

If configuring policies for a fully functional network is difficult, doing so in a way that ensures policy compliance in the face of failures can be almost impossible. Consider the data center network in Figure~\ref{fig:example2} with routers organized as a fat tree and running BGP.\footnote{To scale and to simplify policy implementation, data center networks increasingly use BGP internally, with a private AS number per router~\cite{bgp-in-dc-rfc}.} The network has two clusters, one that hosts services that should be reachable globally and one that hosts that should be accessible only internally. This policy is enabled by using non-overlapping address space in the two clusters and ensuring that only the address space for the global services is announced externally. In addition, to reduce the number of prefixes that are announced externally, the global space is aggregated into one larger, less-specific prefix $P_G$. The semantics of aggregation in BGP is that the aggregate prefix is announced as long as the router has a path to even one sub-prefix.

The operator may decide that a simple way to implement the policy is to have X and Y: $i)$ not export externally what they hear from routers G and H since these routers border the cluster with local services; and $ii)$ announce what they hear from routers C and D and aggregate to $P_G$ if an announcement is subset of $P_G$. The appeal of this implementation is that X and Y do not need to be made aware of which prefixes are global versus local and IP address assignment can be done independently (e.g., new prefixes can be added to local services without updating router configurations).

However, this implementation is incorrect because it does not have the right behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for $P_{l*}$ from C and D, having traversed from G and H to Y to C and D. Per policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services were sensitive) or availability (e.g., if the $P_{l*}$ prefixes are used for other services outside of the data center). This problem does not manifest without failures because then X has and prefers paths to $P_{l*}$ through G and H since they are shorter. A similar problem will happen if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}  
Link failures in data centers are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

\todo{remove if we do not discuss aggregation in the design}
To avoid this problem, the operator may decide to disallow paths with "valleys," i.e., those that go up, down, and back up again. This safeguard can be implemented by $X$ and $Y$ rejecting paths through the other. However, that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{xx}. Suppose links D--A and X--C fail. Now, X will hear announcements for $P_{g2}$ from D and will thus announce the $P_G$ externally. This announcement will bring to X traffic for $P_{g1}$ as well, but because of valley-free filtering, X does not have a valid route for $P_{g1}$ and will thus drop all traffic to it.

Thus, we see that devising configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname helps operators by implementing their high-level policy specification in a way that guarantees compliance under all failures if that is possible. Otherwise, it generates a compile-time error that informs them that the specification cannot be met. For aggregation, it will also provide a lower bound to operators on the number of failures under which aggregation will not result in blackholes.

%\subsection{OLD -- Overview}
%\todo{this subsection can be deleted once we have captured everything in it}
%
%Sources of bugs we fix. (It seems like the main advantage of our approach is not just fixing bugs though - it is easily describing high-level intention).
%Perhaps the best way to do this is by going through a bunch of examples in section 2 and showing how you could easily introduce bugs:
%
%\begin{itemize}
%	\item Out of sync, or copy paste errors due to replicated configs (never an issue due to centralized control)
%	\item Correct filtering to ensure no undesired traffic can flow through the network (e.g., best practices, like an AS should filter customers for their prefix, are implemented automatically )
%	\item Failures can easily lead to unexpected behavior (e.g., a datacenter failure scenario w/instability.)
%	\item Trying to do anything interesting, like get backups correct is difficult (e.g., setting up aggregation wrong)
%	\item Related to the last point, things like aggregation can introduce black holes. We have the information needed to prevent this
%\end{itemize}
%
%
%Possible Examples:
%\begin{itemize}
%	\item Basic datacenter with spine preference
%	\item Simple AS that prefers customers over peers over providers
%	\item Combined internal backup routing with preference based entrance into the network using aggregation.
%	\item Something like cold potato routing
%\end{itemize}
